\sysfull

### 2018-08-05 paper/asplos19/implementation.tex
implemenation: paging

\subsection{\textcolor{red}{to be revised after Virtual Mem.}Paging}
\label{sec:impl:paging}

As described in Section~\ref{sec:memory_virtulaization}, \sys's memory
virtualization scheme uses three buffers: a volatile \texttt{working} buffer, a
non-volatile main memory \texttt{\underline{store}}, and a non-volatile
\texttt{\underline{shadow}} page buffer. 
%
The main memory \texttt{\underline{store}} holds the last committed copy of all 
pages. 
%
After reboot, the \texttt{working} buffer is empty. Execution fills 
the working buffer with pages as the program accesses them. 
%
The program primarily reads and writes data paged into the volatile working memory. 
%
The non-volatile \texttt{\underline{shadow}} buffer serves as an intermediate,
non-volatile backing store for dirty pages evicted from the space-constrained
working buffer.

To access a protected variable with either \texttt{RP}
or \texttt{WP}, \sys searches for the variable in its working buffer first. If
found, the program accesses the data in the working buffer.  If the page is not 
in the working memory, the memory manager fetches it from non-volatile memory. 
%
\sys uses address-based page tagging to make finding a variable efficient.  The
upper bits of a variable's memory address identify its page, and the lower bits
denote the variable's offset in its page. The total number of pages in memory,
$p$, determines the number of tag bits, which is $\log P$.

Page tagging imposes a data alignment requirement.
Page size $S$ must be a power of two. 
%
Pages must be aligned to an $S$-byte boundary. 
%
We experimented with 32-, 64-, 128- and 256\,B pages, an 8\,KB non-volatile \texttt{\underline{shadow}} buffer and
an 8\,KB non-volatile \texttt{\underline{store}}, and a volatile working buffer
of 1\,KB, which is half of the total SRAM size in the MSP430FR6989 microcontroller.
%
\subsection{Coalescing and Commit}
%
\sys uses a two-phase commit process to atomically commit pages despite
possible power interruptions during commit.  When a coalesced task completes,
\sys backs up dirty pages in volatile working memory to the non-volatile
shadow buffer using DMA. 
%
If power fails while backing up these pages, commit fails and the execution
resumes from the last committed point.  After backing up pages from working
memory to the shadow buffer, \sys sets a flag that indicates an on-going atomic
commit.  At this point, commit is irrevocable and will continue until it
completes despite power failures.   
%
Each page is double buffered, with one copy always in the shadow page buffer
and one copy always in the main memory store.  \sys uses double buffering to
commit dirty pages efficiently by swapping which buffer is which, without
copying any data.
%
\subsection{Task Downscaling}
%
\sys's task downscaling preserves forward progress by partially committing an
ongoing, non-terminating task. 
%
Partial commit relies on the existence of a hardware timer, which is a very
common feature in a typical MCU and consumes very little energy (i.e.,
$\approx$3\,$\mu$A/MHz~\cite{msp430datasheet}).

>>>>

\subsection{Paging}
\label{sec:impl:paging}

As described in Section~\ref{sec:memory_virtulaization}, \sys's memory
virtualization scheme uses three buffers: a volatile \texttt{working} buffer, a
non-volatile \texttt{\underline{private}}, and a non-volatile
\texttt{\underline{shadow}} buffer. 
%
The main memory \texttt{\underline{private}} holds the last committed copy of all 
pages. 
%
After reboot, the \texttt{working} buffer is empty. Execution fills 
the \texttt{working} buffer with privatized pages as the program accesses them. 
%
The \texttt{\underline{shadow}} buffer serves as an intermediate,
non-volatile backing \texttt{\underline{private}} for dirty pages evicted from the space-constrained
\texttt{working} buffer.

To access a protected variable with either \texttt{RP}
or \texttt{WP}, \sys searches for the variable in its \texttt{working} buffer first.
 If the page is not, the VMM fetches it from non-volatile memory. 
%
\sys uses address-based page tagging to make finding a variable efficient.  The
upper bits of a variable's memory address identify its page, and the lower bits
denote the variable's offset in its page. The total number of pages in memory,
$p$, determines the number of tag bits, which is $\log P$.

Page tagging imposes a data alignment requirement.
Page size $S$ must be a power of two. 
%
Pages must be aligned to an $S$-byte boundary for efficient memory access. 
%
We experimented with 32-, 64-, 128- and 256\,B pages, an 8\,KB non-volatile \texttt{\underline{shadow}} buffer and
an 8\,KB non-volatile \texttt{\underline{private}}, and a \texttt{working} buffer
of 1\,KB.
%
\subsection{Coalescing and Commit}
%
\sys uses a two-phase commit process to atomically commit pages despite
possible power interruptions during commit.  When a coalesced task completes,
\sys backs up dirty pages in volatile working memory to the non-volatile
shadow buffer using DMA. 
%
If power fails while backing up these pages, commit fails and the execution
resumes from the last committed point.  After backing up pages from working
memory to the shadow buffer, \sys sets a flag that indicates an on-going atomic
commit.  At this point, commit is irrevocable and will continue until it
completes despite power failures.   
%
Each page is double buffered, with one copy always in the shadow page buffer
and one copy always in the main memory store.  \sys uses double buffering to
commit dirty pages efficiently by swapping which buffer is which, without
copying any data.
%
\subsection{Task Downscaling}
%
\sys's task downscaling preserves forward progress by partially committing an
ongoing, non-terminating task. 
%
Partial commit relies on the existence of a hardware timer, which is a very
common feature in a typical MCU and consumes very little energy (i.e.,
$\approx$3\,$\mu$A/MHz~\cite{msp430datasheet}).
